---
title: "Interest Areas for Text Stimuli"
author: "Jenna Duclos"
date: "October 25, 2016"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interest Areas}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Building DataViewer Compatible Interest Area Files

Eye movement data can be analyzed by using interest area definitions
to pick out specific screen areas for summarizing local gaze
behavior. For reading experiments, interest areas may focus on a
single word or a multi-word unit depending on needs of the
analysis. Given an interest area specification, SRR DataViewer can
generate an interest area report to provide information on various
gaze measures for each area: fixation count, summed fixation duration,
regressions into an interest area, etc.

### Using R Package FDBeye to Create SRR Interest Area Files

At this point we assume that the user has created paired bitmap (used
as experimental stimuli in, e.g., SRR Experiment Builder) and comma separated
**region** files. File names for the latter will typically end with
"region.csv".  R and the package FDBeye can be used to create Interest Area
Templates compatible with SRR DataViewer.

The following R packages are required:

* [FDButils](https://github.com/davebraze/FDButils)
* [FDBeye](https://github.com/davebraze/FDBeye)
* [ggplot2](https://cran.r-project.org/web/packages/ggplot2/index.html)
* [readr](https://cran.r-project.org/web/packages/readr/index.html)
* [stringr](https://cran.r-project.org/web/packages/stringr/index.html)
* [yaml](https://cran.r-project.org/web/packages/yaml/index.html)
  (used internally-do not need to load)

#### Steps to get from a `*.region.csv` file to an SRR `*.ias` file.
1. Read the `*.region.csv` file into a data.frame.
2. Use its content to generate a region definition character vector.
3. Write the region definition vector out to file.
4. Hand edit the regdef file to specify the desired regions.
5. Use the regdef file to generate an `*.ias` file suitable for use in
   DataViewer.

Here are the first three steps in R code:

```{r, eval=FALSE}
reg <- read_csv("trial1.region.csv", trim_ws=FALSE)
tmp <- reg2regdef(reg, mrgn.left=86, rgn.maxH=60, rgn.minH=43, rgn.padL=18, rgn.padR=18)
outf <- "trial1.region.def"
writeLines(tmp, con=outf)
```

The region definition file will typically end with
`*.region.def`. Once the file is saved to the working directory it can
be hand edited in any text editor. The text document will include a
block of information text layout on the screen as seen by the study
participants. This appears in a 'yaml' block at the beginning of the
file. Some of this information is read directly from the `region.csv`
file, and some is specified in the function call `reg2regdef()`. It is
easiest to specify this information when the region definition file is
created, rather than hand editing it in the `region.def` file.

The text you want to edit to specify interest area locations will be
at the bottom of the `*.region.def` file and will look something like
this:

```
The quick brown fox jumps over the lazy dog.
...|.....|.....|...|.....|....|...|....|....
```

Each line of a text stimulus is paired with a 'regioning
string'. Within a regioning string, each letter and punctuation mark
in the stimulus sentence is marked with a dot, "**.**", and regions
are separated with pipe character, "**|**". The pipe character marks
the first character of a new region. In this example, all regions in
the stimulus sentence but the first will begin with a "space". (The
first region of a line always begins with the first character.) To
edit locations of the interest areas, add or remove "**|**" characters
from the regioning string. Note that the regioning string must be the
same length as its paired text line. To maintain the same length
between text line and regioning string, any removed "**|**" must be
replaced with a "**.**".

An edited regioning string in the text document will look like this:

```
The quick brown fox jumps over the lazy dog.
...................|.....|........|.........
```

Once you have hand edited the text document to reflect the interest
area locations you want save the file.

#### Preview the new interest areas:
1. Use the `*.region.def` file to build a set of regions.
2. Load the bitmap corresponding to the regions.
3. Plot the new interest areas over the bitmap.

Here are the steps in R code:

```{r, eval=FALSE}
ias2 <- regdef2ias("trial1.region2.def")
write_delim(ias2, "trial1.region2.ias", "\t", col_names=FALSE)
fp2 <- fixPlot(data = data.frame(x=-1, y=-1),
              bgImage = "trial1.png", bgAlpha=1,
              xyMap = ggplot2::aes_string(x='x', y='y'),
              pointMap=ggplot2::aes_string(alpha=0)
              )
fp2 + geom_rect(data=ias2, inherit.aes=FALSE,
               aes(xmin=x1+1, xmax=x2, ymin=y1, ymax=y2,
                   fill=NULL, color=as.factor(regnum%%2)), alpha=0) +
     guides(color=FALSE)
```
Note that for our purposes, we renamed the `*region.def` file to
reflect the changes made (`*region.def` > `*region2.def`). We chose to
save and document the process and all the steps included in generating
the new interest areas. Unless specified in your own code to be saved
with a new name, the original `*region.def` and `*.ias` files will be
overwritten.

Once the bitmaps and interest area regions have been displayed in R
Plots you can visualize the regions as they will appear in
DataViewer. If the text regions are not what you want/expect, then you
can re-edit your `*region.def` file, and re-run regdef2ias(), and
reload the plot. If the top and bottom boundaries of the regions are
not placed appropriately, they can be adjusted by changing argument
values rgn.maxH, and rgn.minH to `reg2regdef()`. See help page for
that function.

```
reg2regdef(reg, mrgn.left=86, rgn.maxH=60, rgn.minH=43, rgn.padL=18, rgn.padR=18)
```

It is recommended that one `*region.def` file be hand edited and one
`*.ias` file be generated and checked for accuracy before creating
multiple files. Then when you are sure the interest areas locations
are appearing as you would like you can generate a full set of region
definition files to edit.

#### Create all the `*region.def` and `*.ias` files you need:
1. Set your working directory to the folder containing all of the
   `*region.csv`.
2. Make a vector of all the files located in your working directory
   and remove all files that are not `*region.csv`.
3. Using a loop function, create one `*region.def` for each
   `*region.csv`.
4. Hand edit the regioning string in each `*region.def`.
5. Then return to R and make a vector of all the files located in your
   working directory and remove all files that are not `*region.def`.
6. Using a loop function, create one `*.ias` for each `*region.def`.

Here are steps 2-3 in R code:

```{r, eval=FALSE}
flist <- dir()
flist <- flist[str_detect(flist, "region[.]csv")] ## change the regexp as needed to pick out your files
for (ff in flist) {
    reg <- read_csv(ff, trim_ws=FALSE)
    rdef <- reg2regdef(reg, mrgn.left=86, rgn.maxH=60, rgn.minH=43, rgn.padL=18, rgn.padR=18)
    outf <- str_replace(ff, "region.csv$", "region2.def")
    writeLines(rdef, con=outf)
}
```

Here are steps 5-6 in R code:

```{r, eval=FALSE}
flist <- dir()
flist <- flist[str_detect(flist, "region2[.]def")]
for (ff in flist) {
    ias <- regdef2ias(ff)
    outf <- str_replace(ff, "def$", "ias")
    write_delim(ias, outf, delim="\t", col_names=FALSE)
}
```

Now that your interest area files have been created for each trial you will need to load them into DataViewer. For theses steps refer to **Interest Areas for Data Analysis**.
